---
title: "Seabird Sightings"
output:
  html_document:
    df_print: paged
---

```{r message = F, warning = F}
library(here)
library(tidyverse)
```
## Seabirds data

The seabirds data contains sightings of seabirds from 1969 - 1990. 
Each observation includes the species common name, scientific name, abbreviated name, number of individuals per sighting, date and latitude. 

### Data was cleaned in `cleaning_script.R` to achieve the following:
* Change variable names to short, meaningful names
* Join bird data to ship data to retrieve locations of sightings
* Remove NAs in count and species abbreviations so that each observation indicates a sighting
* Remove information on age and plumage from species names variables 
* Verify values for count and latitude are within specified range
* Export clean data to a new csv file within the clean_data folder, named `seabirds_cleaned.csv`

```{r message = F, warning = F}
seabirds_full_clean <- read_csv(here("clean_data/seabirds_cleaned.csv"))
```

<br>

#### Bird with the most individual sightings
```{r message = F, warning = F}
# Make a table of common, scientific and abbreviated species names
seabirds_species <- seabirds_full_clean %>% 
  select(species_scientific_name,
         species_abbreviation,
         species_common_name) %>%
  distinct()
```


```{r message = F, warning = F}
seabirds_full_clean %>%
  group_by(species_abbreviation) %>%
  summarise(sightings_count = n()) %>%
  slice_max(sightings_count, n=1) %>%
  left_join(seabirds_species, by = "species_abbreviation")
```
The bird with the most sightings was the Wandering albatross, with 11,247 sightings. 

<br>

#### Bird with the highest total count

```{r message = F, warning = F}
seabirds_full_clean %>%
  group_by(species_abbreviation) %>%
  summarise(sightings_count = n(),
            total_count = sum(count)) %>%
  slice_max(total_count, n=1) %>%
  left_join(seabirds_species, by = "species_abbreviation")

```
The bird with the highest total count was the Short-tailed shearwater. 

<br>

#### Bird with the highest total count above latitude -30
```{r message = F, warning = F}
seabirds_full_clean %>%
  filter(lat > -30) %>%
  group_by(species_abbreviation) %>%
  summarise(sightings_count = n(),
            total_count = sum(count)) %>%
  slice_max(total_count, n=1) %>%
  left_join(seabirds_species, by = "species_abbreviation")
```
The bird with the highest total count in latitudes above -30 was the Wedge-tailed Shearwater. 

<br>

#### How many different types of birds only ever seen in groups of 1

```{r message = F, warning = F}
#table of birds sighted once
seabirds_single_sightings <- seabirds_full_clean %>%
  filter(count == 1) %>%
  group_by(species_abbreviation, genus) %>%
  summarise(sightings = n())

#table of birds sighted in groups
seabirds_group_sightings <- seabirds_full_clean %>%
  filter(count > 1) %>%
  group_by(species_abbreviation) %>%
  summarise(sightings = n())

#perform anti-join to find birds who were not sighted in a group
anti_join(seabirds_single_sightings,
          seabirds_group_sightings,
          by = "species_abbreviation") %>%
  nrow()

```

There are 23 species of birds that are only ever seen individually and not in groups.


```{r message = F, warning = F}
#perform anti-join to find birds who were not sighted in a group
anti_join(seabirds_single_sightings,
          seabirds_group_sightings,
          by = "genus") %>%
  nrow()
```

There are 6 genera of bird whose species were only ever seen individually

<br>

#### How many penguins were seen

```{r message = F, warning = F}
seabirds_full_clean %>%
  filter(str_detect(species_common_name, "[Pp]enguin")) %>%
  summarise(total_penguin_count = sum(count),
            total_genus_count = length(unique(genus)),
            total_penguin_species_count = length(unique(species_common_name)))
```

A total of 158 penguins were sighted consisting of 6 different genera and 7 species.