---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---
```{r message = F, warning = F}
library(tidyverse)
library(here)
```

# Candy Data


The halloween candy dataset are the results of an online survey of trick or treat halloween candy hauls. 
Each participant responds to each "candy" (or other item) with a feeling of joy, despair or meh (indifference) if they were to find said item in their halloween haul. 
Participants were also able to enter their age, gender, whether or not they were actually trick or treating and their country. 

## Data Cleaning


Before running analysis, `candy_data_cleaning.R` must be run in order to export the clean data as a csv. 
Steps taken to clean the data:

* merging datasets from 2015, 2016 and 2017 and ensuring that column names were consistent
* country & age variables cleaned for consistency
* columns containing "joy/despair/meh" were interpreted to be candy columns
* each participant received a unique id
* table was pivoted to convert candy columns into rows
* column is_candy indicates if the item is indeed a candy


```{r message = F, warning = F}
#due to the number of NAs in country and gender, need to specify col type
candy_data <- read_csv(here::here("clean_data/candy_clean.csv"), 
                  col_types = cols(country = col_character(),
                                   gender = col_character()
                                   ))
```

## Data Analysis 


### What is the total number of candy ratings given across the three years

```{r message = F, warning = F}
candy_nrow <- candy_data %>%
  filter(is_candy == TRUE) %>%
  nrow()
```

The total number of ratings across all items is `r nrow(candy_data)`. Of those approximately `r candy_nrow` are candy.


### What was the average age of people who are going out trick or treating and the average age of people not going trick or treating?

```{r message = F, warning = F}
candy_data %>%
  group_by(going_out) %>%
  summarise(avg_age = mean(age, na.rm = TRUE))
```

### For each of joy, despair and meh, which candy bar received the most of these ratings?

```{r message = F, warning = F}
# For items that are candy 
candy_data %>%
  filter(is_candy == TRUE) %>%
  group_by(response, candy) %>%
  summarise(num_ratings = n()) %>%
  group_by(response) %>%
  slice_max(num_ratings)
```

```{r message = F, warning = F}
# For all items surveyed
candy_data %>%
  group_by(response, candy) %>%
  summarise(num_ratings = n()) %>%
  group_by(response) %>%
  slice_max(num_ratings)
```

### How many people rated Starburst as despair?

```{r message = F, warning = F}
candy_data %>% 
  filter(candy == "starburst" & response == "DESPAIR") %>%
  count()
```

## Candy Ratings

For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r message = F, warning = F}
candy_ratings <- candy_data %>% 
  mutate(rating_score = 
          case_when(
            response == "JOY" ~ 1,
            response == "DESPAIR" ~ -1,
            response == "MEH" ~ 0
          ))
```


### What was the most popular candy bar by this rating system for each gender in the dataset?

```{r message = F, warning = F}
candy_ratings %>%
  filter(is_candy == TRUE) %>%
  filter(!is.na(gender)) %>%
  group_by(candy, gender) %>%
  summarise(avg_rating = mean(rating_score)) %>%
  group_by(gender)%>%
  slice_max(avg_rating, n = 2)
```
Any full sized candy bar is the most popular accross all genders. The next most popular is peanut butter cups for both males and females, twix for other and kit_kats for those who'd rather not say. 



```{r message = F, warning = F}
candy_ratings %>%
  filter(!is.na(gender)) %>%
  group_by(candy, gender) %>%
  summarise(avg_rating = mean(rating_score)) %>%
  group_by(gender)%>%
  slice_max(avg_rating, n = 2)
```
If we ignore any full sized candy bars, and look at all other items rated (including non-candy), the next most popular item for the male category is still a peanut butter cup. For all other gender categories, it is cash or other forms of legal tender. 


### What was the most popular candy bar in each year?
```{r message = F, warning = F}
candy_ratings %>% 
  filter(candy != "any_full_sized_candy_bar" & is_candy == TRUE) %>%
  group_by(candy, year) %>%
  summarise(avg_rating = mean(rating_score)) %>%
  group_by(year)%>%
  slice_max(avg_rating)
```
Across all years, any full sized candy bar takes top place. Next in line is peanut butter cups in 2015 and 2017. Kit Kats were the next best thing after a full sized bar in 2016. 


### What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
```{r message = F, warning = F}
candy_ratings %>%
  filter(is_candy == TRUE) %>%
  mutate(country_group = recode(country,
    "usa" = "usa",
    "uk" = "uk",
    "canada" = "canada",
    .default =  "other"
  )) %>%
  group_by(candy,country_group) %>%
  summarise(rating = mean(rating_score)) %>%
  group_by(country_group) %>%
  slice_max(rating)
```
Any full sized candy bar takes top place in all countries. In the UK it shares top place with rolos. 

```{r message = F, warning = F}
candy_ratings %>%
  mutate(country_group = recode(country,
    "usa" = "usa",
    "uk" = "uk",
    "canada" = "canada",
    .default =  "other"
  )) %>%
  group_by(candy,country_group) %>%
  summarise(rating = mean(rating_score)) %>%
  group_by(country_group) %>%
  slice_max(rating)
```
If we look at all items, including non-candy items, the UK rates cash the higher than any full sized candy bar. 

