---
title: "R Notebook"
output: html_notebook
---
```{r}
library(tidyverse)
library(here)
```

```{r}
candy_data <- read_csv(here::here("clean_data/candy_clean.csv"), 
                  col_types = cols(country = col_character(),
                                   gender = col_character()
                                   ))
```

What is the total number of candy ratings given across the three years. (number of candy ratings, not number of raters. Don’t count missing values)

```{r}
nrow(candy_data)
```

What was the average age of people who are going out trick or treating and the average age of people 3. not going trick or treating?

```{r}
candy_data %>%
  group_by(going_out) %>%
  summarise(avg_age = mean(age, na.rm = TRUE))
```
For each of joy, despair and meh, which candy bar revived the most of these ratings?

```{r}
candy_data %>%
  group_by(response, candy) %>%
  summarise(num_ratings = n()) %>%
  group_by(response) %>%
  slice_max(num_ratings)

```
How many people rated Starburst as despair?

```{r}
candy_data %>% 
  filter(candy == "starburst" & response == "DESPAIR") %>%
  count()
```

For the next three questions, count despair as -1, joy as +1 and meh as 0.
```{r}
candy_ratings <- candy_data %>% 
  mutate(rating_score = 
          case_when(
            response == "JOY" ~ 1,
            response == "DESPAIR" ~ -1,
            response == "MEH" ~ 0
          ))
```


What was the most popular candy bar by this rating system for each gender in the dataset?

```{r}
candy_ratings %>%
  filter(!is.na(gender)) %>%
  group_by(candy, gender) %>%
  summarise(rating = sum(rating_score)) %>%
  group_by(gender)%>%
  slice_max(rating)
```

What was the most popular candy bar in each year?
```{r}
candy_ratings %>% 
  group_by(candy, year) %>%
  summarise(rating = sum(rating_score)) %>%
  group_by(year)%>%
  slice_max(rating)
```

What was the most popular candy bar by this rating for people in US, Canada, UK and all other countries?
```{r}
candy_ratings %>%
  mutate(country_group = case_when(
    country == "usa" ~ "usa",
    country == "uk" ~ "uk",
    country == "canada" ~ "canada",
    TRUE ~ "other"
  )) %>%
  group_by(candy,country_group) %>%
  summarise(rating = sum(rating_score)) %>%
  group_by(country_group) %>%
  slice_max(rating)
```

